{Simulating the Brownian Dynamics of a Two-dimensional Model for the Dynein Motor Protein}\vspace{1.5em}\\ \includegraphics[width=0.5\columnwidth]{OSU_logo.png}\vspace{0.5em}\\ \large{Oregon State University}

### 2019-04-23 papers/waczak-thesis/thesis/chapters/dynein.tex
big thesis update 1

Dynein walks along a particular cyctoskeleton filament called a microtubule. Microtubules are cytoskeletal structures formed by self assembly of $\alpha$ and $\beta$ tubulin heterodimers (shown as green in Figure \ref{fig:cytoskeleton}). \cite{downing_tubulin_1998} These molecular structures are typically comprised of 13 protofilaments each of which is a periodic arrangement of the $\alpha$ and $\beta$ dimers.

Microtubules exhibit an inherent polarity identified as the plus ($+$) and minus ($-$) ends \cite{lodish_microtubule_2000}. This polarity is both function and structural. One end of the microtubule experiences much faster addition of tubulin making it easy to identify\cite{heidemann_visualization_1980}. For our purposes though, the structural polarity of the microtubule is more important as the formation of the microtubule results in a helical structure that has definite chirality \cite{microtubule_chirality}. Two possible microtubule configurations are shown in Figure \ref{fig:chirality}. Note that even though there is a deviation by angle $\theta$ between the two microtubules, the overall handedness of the helix is the same.

@@

As Figure \ref{fig:cytoskeleton} shows, microtubules foliate the entire cell with paths that enable motor proteins to move cargo. Most(not all) variations of the kinesin molecular motor are plus-end directed. They are almost always observed to walk towards the plus-end of the microtubule. Dynein is strictly a minus-end directed motor \cite{alberts_molecular_2002}. How this directed motion is derived from the symmetric dynein structure is still an open question.\\

\section{The powerstroke model} 
Take another look at Figure \ref{fig:motor proteins}. How does this complicated structure perform complex locomotion using only the resources of its aqueous environment? How can a highly symmetric structure lead to directed motion? The chemical resolution to this problem arises from the interaction of ATP with each of the motor domains of dynein. The cycle of converting the chemical energy of  ATP into useful mechanical energy for motion is called the \textit{mechanochemical cycle}\cite{cianfrocco_mechanism_2015}. Describing all of the chemical interactions involved in a single dynein step has been an active area of research. The currently accepted model was proposed by Cianfrocco et al. and is summarized in the following graphic shown in Figure \ref{fig:mechanochemical_cycle}.

\begin{figure}[!hbt]
	\centering
	\includegraphics[width=1\columnwidth]{mechanochemical_cycle}
	\caption{The mechanochemical cycle of dynein. Dynein goes through many identifiable stages as ATP from the environment interacts with the motor domain and changes the conformation of the protein. Image taken from \cite{cianfrocco_mechanism_2015}.}
	\label{fig:mechanochemical_cycle}
\end{figure}

@@

\begin{figure}[!hbt]
	\centering
	\includegraphics[width=1\columnwidth]{pre_post_powerstroke}
	\caption{An illustration of the pre and post stroke states of dynein in relation to the microtubule and the cargo that they carry. This two-dimensional illustration shows only one of legs of dynein. Image taken from \cite{dynein_powerstroke}.}
\end{figure}

Scientists have determined the steps of this chemical cycle by carefully isolating and analyzing the resulting structures via cryo-electron microscopy \cite{cryo_em}. While this is useful for determining the conformation before and after a chemical process occurs, it does note describe the actual motion along the microtubule-- only intermittent states. All of the information about how flexible each domain is during binding and to what extent the protein diffuses in its search for the microtubule require alternative methods to establish.

\section{Other observational techniques} 
 Kinesin and Dynein have masses of a few hundred kDa and  $\sim2$ MDa respectively \cite{liao_kinesin_1998, johnson_structure_1983}. Recalling that 1 Dalton (Da) corresponds to one atomic mass unit means these proteins consist of, at most, a few million atoms. This small scale poses a unique experimental challenge. One way this is approached at Prof. Weihong Qiu's group is by attaching small fluorescent molecules called fluorophores to the binding sites that motor proteins use to latch onto their cargo. Then, the individual movement of motors can be tracked using total internal reflection fluorescence microscopy to produce video of the proteins in action as illustrated in figure \ref{fig:weihong tirf} from \cite{qiu_dynein_2012}.

\begin{figure}[!hbt]
	\centering
	\includegraphics[angle=90, width=0.75\columnwidth]{weihong_motor}
	\caption{\textbf{TIRF Microscopy Imaging} individual kinesin proteins are imaged as small green dots and tracked as they move. Slices at different times are stacked so that the slope of each green line indicates the velocity of the protein. Image taken from \cite{qiu_dynein_2012}.}
	\label{fig:weihong tirf}
\end{figure}

Figure \ref{fig:weihong tirf} demonstrates fluorescence microscopy enables observation of protein speed, how far it walks along the microtubule, which direction is preferred. Furthermore, by tagging a motor domain instead of the tail, this technique can resolve step sizes and as shown in Figure \ref{fig:weihong histogram}. \cite{qiu_dynein_2012} Taking into account these studies as well as the delicate cryo-em analysis forms a rough sketch for dynein's motion. Unfortunately though, this is the limit of current experimental methods; observing the individual motions of each of the binding domains of a dynein protein is not yet possible.

>>>>

Dynein walks along a particular cyctoskeleton filament called a microtubule. Microtubules are cytoskeletal structures formed by self assembly of $\alpha$ and $\beta$ tubulin heterodimers (shown as green in Figure \ref{fig:cytoskeleton}). \cite{downing1998tubulin} These molecular structures are typically comprised of 13 protofilaments each of which is a periodic arrangement of the $\alpha$ and $\beta$ dimers.

Microtubules exhibit an inherent polarity identified as the plus ($+$) and minus ($-$) ends \cite{berk2000molecular}. This polarity is both function and structural. One end of the microtubule experiences much faster addition of tubulin making it easy to identify\cite{heidemann1980visualization}. For our purposes though, the structural polarity of the microtubule is more important as the formation of the microtubule results in a helical structure that has definite chirality \cite{hunyadi2007microtubule}. Two possible microtubule configurations are shown in Figure \ref{fig:chirality}. Note that even though there is a deviation by angle $\theta$ between the two microtubules, the overall handedness of the helix is the same.

@@

As Figure \ref{fig:cytoskeleton} shows, microtubules foliate the entire cell with paths that enable motor proteins to move cargo. Most(not all) variations of the kinesin molecular motor are plus-end directed. They are almost always observed to walk towards the plus-end of the microtubule. Dynein is strictly a minus-end directed motor \cite{alberts2002molecular}. How this directed motion is derived from the symmetric dynein structure is still an open question.\\

\section{The powerstroke model} 
Take another look at Figure \ref{fig:motor proteins}. How does this complicated structure perform complex locomotion using only the resources of its aqueous environment? How can a highly symmetric structure lead to directed motion? The chemical resolution to this problem arises from the interaction of ATP with each of the motor domains of dynein. The cycle of converting the chemical energy of  ATP into useful mechanical energy for motion is called the \textit{mechanochemical cycle}\cite{cianfrocco2015mechanism}. Describing all of the chemical interactions involved in a single dynein step has been an active area of research. The currently accepted model was proposed by Cianfrocco et al. and is summarized in the following graphic shown in Figure \ref{fig:mechanochemical_cycle}.

\begin{figure}[!hbt]
	\centering
	\includegraphics[width=1\columnwidth]{mechanochemical_cycle}
	\caption{The mechanochemical cycle of dynein. Dynein goes through many identifiable stages as ATP from the environment interacts with the motor domain and changes the conformation of the protein. Image taken from \cite{cianfrocco2015mechanism}.}
	\label{fig:mechanochemical_cycle}
\end{figure}

@@

\begin{figure}[!hbt]
	\centering
	\includegraphics[width=1\columnwidth]{pre_post_powerstroke}
	\caption{An illustration of the pre and post stroke states of dynein in relation to the microtubule and the cargo that they carry. This two-dimensional illustration shows only one of legs of dynein. Image taken from \cite{lin2014structural}.}
\end{figure}

Scientists have determined the steps of this chemical cycle by carefully isolating and analyzing the resulting structures via cryo-electron microscopy \cite{burgess2003dynein}. While this is useful for determining the conformation before and after a chemical process occurs, it does not describe the actual motion along the microtubule-- only intermittent states. All of the information about how flexible each domain is during binding and to what extent the protein diffuses in its search for the microtubule require alternative methods to establish.

\section{Other observational techniques} 
 Kinesin and Dynein have masses of a few hundred kDa and  $\sim2$ MDa respectively \cite{liao1998kinesin, johnson1983structure}. Recalling that 1 Dalton (Da) corresponds to one atomic mass unit means these proteins consist of, at most, a few million atoms. This small scale poses a unique experimental challenge. One way this is approached at Prof. Weihong Qiu's group is by attaching small fluorescent molecules called fluorophores to the binding sites that motor proteins use to latch onto their cargo. Then, the individual movement of motors can be tracked using total internal reflection fluorescence microscopy to produce video of the proteins in action as illustrated in figure \ref{fig:weihong tirf} from \cite{qiu2012dynein}.

\begin{figure}[!hbt]
	\centering
	\includegraphics[angle=90, width=0.75\columnwidth]{weihong_motor}
	\caption{\textbf{TIRF Microscopy Imaging} individual kinesin proteins are imaged as small green dots and tracked as they move. Slices at different times are stacked so that the slope of each green line indicates the velocity of the protein. Image taken from \cite{qiu2012dynein}.}
	\label{fig:weihong tirf}
\end{figure}

Figure \ref{fig:weihong tirf} demonstrates fluorescence microscopy enables observation of protein speed, how far it walks along the microtubule, which direction is preferred. Furthermore, by tagging a motor domain instead of the tail, this technique can resolve step sizes and as shown in Figure \ref{fig:weihong histogram}. \cite{qiu2012dynein} Taking into account these studies as well as the delicate cryo-em analysis forms a rough sketch for dynein's motion. Unfortunately though, this is the limit of current experimental methods; observing the individual motions of each of the binding domains of a dynein protein is not yet possible.

### 2019-04-23 papers/waczak-thesis/thesis/chapters/introduction.tex
big thesis update 1

Cells are the fundamental unit for all life on Earth. In order
for them to perform their biological tasks, the complex structures they contain
must be able to replicate and move within their aqueous environment. This
transport problem  poses a unique challenge as diffusion alone is far to slow,
and random, a process to be able to produce movement on cellular timescales[].
Nature's solution to this problem is twofold. Within each cell there is a vast network of interweaving filaments called the \textit{cytoskeleton} that spans the gaps between the substructures of the cell. This cytoskeleton serves as a highway to a class of proteins called \textit{molecular motors} which can be distinguished into three types: kinesin, myosin, and dynein. Molecular motors attach to cellular cargo and drag it along the cytoskeleton by hydrolyzing adenosine triphosphate (ATP) in order to convert stored chemical energy into mechanical work\cite{alberts_molecular_2002}. Correct operation of these motors results in processes such as cell division, muscle contraction, and axonal transport; failure can lead to debilitating diseases such as ALS \cite{boillee_als:_2006}.

@@

Astonishingly, each of these motors is composed of two legs that combine to form a bipedal structure (see Figure \ref{fig:motor proteins}). that steps like a human as it moves along the cytoskeleton. Myosin and kinesin behave like a normal person; their steps alternate between opposing domains. This motion is called \textit{hand over hand} stepping \cite{noauthor_drunk_nodate}. Dynein, however, is unique in its motion as it takes hand-over-hand steps and also \textit{inchworms}. This alternative behavior is a scenario in which one motor consistently leads the opposite resulting in a shuffling motion. Furthermore, dynein even takes frequent backwards steps as shown in the stepping histogram in figure \ref{fig:weihong histogram} \cite{qiu_dynein_2012}. Collectively this behavior has earned dynein the title of \textit{drunken walker} which compactly summarizes its peculiar tendencies \cite{noauthor_drunk_nodate}. \\

\begin{figure}[!hbt]
	\centering
	\includegraphics[width=0.5\columnwidth]{weihong_histogram}
	\caption{Stepping histogram showing the distribution of step sizes for \textit{in vitro} dynein. 1-D data corresponds to the projection of the step along the axis of the microtubule. 2-D data corresponds to the total step size over the two-dimensional structure. Figure taken form \cite{qiu_dynein_2012}.}  
	\label{fig:weihong histogram} 
\end{figure}

This unique stepping behavior makes dynein an interesting subject to study. There have been significant efforts aimed at characterizing the complex, multi-phase chemical cycle involved in the hydrolysis of ATP\cite{cianfrocco_mechanism_2015}. However, the dynamics of dynein's physical motion through space remains an open question as current optical techniques have difficulty resolving the individual motion of each dimer. In this thesis, we examine a simplified physical model for the motion of dynein using Brownian dynamics. By simulating this model, we demonstrate it's ability to replicate many of the observed properties of the protein. We will also use the model to comment a proposed mechanism for how dynein generates its movement, namely the powerstroke model.

>>>>

Cells are the fundamental unit for all life on Earth. In order
for them to perform their biological tasks, the complex structures they contain
must be able to replicate and move within their aqueous environment. This
transport problem  poses a unique challenge as diffusion alone is far to slow,
and random, a process to be able to produce movement on cellular timescales[].
Nature's solution to this problem is twofold. Within each cell there is a vast network of interweaving filaments called the \textit{cytoskeleton} that spans the gaps between the substructures of the cell. This cytoskeleton serves as a highway to a class of proteins called \textit{molecular motors} which can be distinguished into three types: kinesin, myosin, and dynein. Molecular motors attach to cellular cargo and drag it along the cytoskeleton by hydrolyzing adenosine triphosphate (ATP) in order to convert stored chemical energy into mechanical work\cite{alberts2002molecular}. Correct operation of these motors results in processes such as cell division, muscle contraction, and axonal transport; failure can lead to debilitating diseases such as ALS \cite{boillee2006disease}.

@@

Astonishingly, each of these motors is composed of two legs that combine to form a bipedal structure (see Figure \ref{fig:motor proteins}). that steps like a human as it moves along the cytoskeleton. Myosin and kinesin behave like a normal person; their steps alternate between opposing domains. This motion is called \textit{hand over hand} stepping \cite{goldman2009drunk}. Dynein, however, is unique in its motion as it takes hand-over-hand steps and also \textit{inchworms}. This alternative behavior is a scenario in which one motor consistently leads the opposite resulting in a shuffling motion. Furthermore, dynein even takes frequent backwards steps as shown in the stepping histogram in figure \ref{fig:weihong histogram} \cite{qiu2012dynein}. Collectively this behavior has earned dynein the title of \textit{drunken walker} which compactly summarizes its peculiar tendencies \cite{goldman2009drunk}. \\

\begin{figure}[!hbt]
	\centering
	\includegraphics[width=0.5\columnwidth]{weihong_histogram}
	\caption{Stepping histogram showing the distribution of step sizes for \textit{in vitro} dynein. 1-D data corresponds to the projection of the step along the axis of the microtubule. 2-D data corresponds to the total step size over the two-dimensional structure. Figure taken form \cite{qiu2012dynein}.}  
	\label{fig:weihong histogram} 
\end{figure}

This unique stepping behavior makes dynein an interesting subject to study. There have been significant efforts aimed at characterizing the complex, multi-phase chemical cycle involved in the hydrolysis of ATP\cite{cianfrocco2015mechanism}. However, the dynamics of dynein's physical motion through space remains an open question as current optical techniques have difficulty resolving the individual motion of each dimer. In this thesis, we examine a simplified physical model for the motion of dynein using Brownian dynamics. By simulating this model, we demonstrate it's ability to replicate many of the observed properties of the protein. We will also use the model to comment a proposed mechanism for how dynein generates its movement, namely the powerstroke model.

### 2019-04-23 papers/waczak-thesis/thesis/chapters/methods.tex
big thesis update
big thesis update 1

>>>>

	
\section{The Code}
	\section{Time evolution: Euler's method}
	\subsection{The Random Force}
\section{Determining constants of the model}
The stalk and tail lengths $\ell_S, \ell_T$ as well as a set of equilibrium angles $\{\varphi_{i,\text{eq}}\}$ were determined by examining microscopy images of dynein in various configurations. 
\begin{figure}[hbt!]
	\centering
	\includegraphics[width=0.3\columnwidth]{schematic-2-superimposed}
	\caption{Our model superimposed over microscopy image of Dynein \textbf{NOTE: add source} for determining lengths and equilibrium angles} 
	\label{fig:superimpmosed}
\end{figure}

### 2019-04-23 papers/waczak-thesis/thesis/chapters/theory.tex
big thesis update 1

Now that I have established some facts about motor proteins as well as specifics about dynein's many odd behaviors, we may begin our discussion of how to construct and test a physical model for its motion. First, we note that dynein's large mass of roughly 2MDa \cite{johnson_structure_1983} as compared to similar motor proteins such as Kinesin (a few hundred kDa) \cite{liao_kinesin_1998} exaggerates the effect of interactions with the  aqueous environment of the cell. This results in a much more significant drag force as well as a random element from water molecule collisions that changes the overall dynamics of the system. The set of equations governing this motion is called Brownian dynamics and is what we use to simulate the trajectory of dynein along the microtubule. 
	\section{Brownian dynamics}
	To model dynein we treat its various domains (binding, motor, and tail) as spheres held together by massless rigid-rods. In general, the dynamics of a constant mass particle are governed by Newton's second law
	\begin{equation}
		\sum_i \mathbf{F}_i = \dot{\mathbf{p}} = m\ddot{\mathbf{x}}
	\end{equation}
	Thus, determining the dynamics of a system boils down to describing the various forces which act upon it and then solving the resulting differential equation for some set of initial conditions. For small particles in a solvent, there are three important forces to consider. The first is linear drag, which we write as 
	\begin{equation}
		F_\text{drag} = -\gamma\dot{\mathbf{x}}
	\end{equation}
	so that the drag force is proportional to the velocity by a factor $\gamma$. For a spherical particle of radius $a$ moving through a solvent with liquid viscosity $\eta$ this was derived by Stoke to be 
	\begin{equation}
		\gamma = 6\pi\eta a
	\end{equation}

@@

	To do this we steer each domain towards an equilibrium configuration by a Hooke's law restorative torque with signed magnitude of the following form
	\begin{equation}
		\Gamma_i = -c_i\left(\varphi_i-\varphi_{i,\text{eq}}\right)
	\end{equation}
	Here the $c_i$ are spring constants (a free parameter of the model) and the set $\{\varphi_i\}$ are the angles measured between the $i-1$ and $i+1$ domains. Equally useful is the harmonic energy associated with each torque. These are given by
	\begin{equation}
		U_i = \frac{1}{2}c_i(\varphi_i-\varphi_{i,\text{eq}})^2
	\end{equation}

@@

	\section{Equations of Motion}
	The dynamics of each domain are determined by equation 3.5 so that in order to simulate the complicated structure we must solve a system of coupled first order differential equations. In doing so, it is convenient to introduce two states to distinguish between dynein's stepping and when both of its binding domains are attached to the microtubule. We shall refer to these configurations hereafter as the one bound state and the both bound state with the idea being that they directly correspond to the pre and post-stroke states defined by the mechanochemical cycle of \cite{cianfrocco_mechanism_2015}.

@@

		\subsection{The one bound state}
		Figure \ref{fig:one bound} illustrates dynein in the one bound state with new angles $\{\theta_i\}$ chosen for convenience of calculation and computation.\\
		\begin{figure}[!hbt]
			\centering
			\includegraphics[angle=-90,width=0.75\columnwidth]{onebound}
			\caption{The dynein model in the one bound state }
			\label{fig:one bound}  
		\end{figure}

@@

		\subsection{The both bound state}

		\section{Determining constants of the model}
		The stalk and tail lengths $\ell_S, \ell_T$ as well as a set of equilibrium angles $\{\varphi_{i,\text{eq}}\}$ were determined by examining microscopy images of dynein in various configurations. 
		\begin{figure}[hbt!]
			\centering
			\includegraphics[width=0.3\columnwidth]{schematic-2-superimposed}
			\caption{Our model superimposed over microscopy image of Dynein \textbf{NOTE: add source} for determining lengths and equilibrium angles} 
			\label{fig:superimpmosed}
		\end{figure}

@@

>>>>

Now that I have established some facts about motor proteins as well as specifics about dynein's many odd behaviors, we may begin our discussion of how to construct and test a physical model for its motion. First, we note that dynein's large mass of roughly 2MDa \cite{johnson1983structure} as compared to similar motor proteins such as Kinesin (a few hundred kDa) \cite{liao1998kinesin} amplifies the effect of interactions with the  aqueous environment of the cell. This results in a much more significant drag force as well as a random element from water molecule collisions that changes the overall dynamics of the system. The set of equations governing this motion is called Brownian dynamics and is what we use to simulate the trajectory of dynein along the microtubule. 
	\section{Brownian dynamics}
	To model dynein we treat its various domains (binding, motor, and tail) as spheres held together by massless rigid-rods. In general, the dynamics of a constant mass particle are governed by Newton's second law
	\begin{equation}
		\sum_i \mathbf{F}_i = \dot{\mathbf{p}} = m\ddot{\mathbf{x}}
	\end{equation}
	Thus, determining the dynamics of a system boils down to describing the various forces which act upon it and then solving the resulting differential equation for some set of initial conditions. For small particles in a solvent, there are three important forces to consider. The first is linear drag, which we write as 
	\begin{equation}
		F_\text{drag} = -\gamma\dot{\mathbf{x}}
	\end{equation}
	so that the drag force is proportional to the velocity by a factor $\gamma$. For a spherical particle of radius $a$ moving through a solvent with liquid viscosity $\eta$ this was derived by Stoke to be 
	\begin{equation}
		\gamma = 6\pi\eta a
	\end{equation}

@@

	To do this we steer each domain towards an equilibrium configuration by a Hooke's law restorative torque with signed magnitude of the following form
	\begin{equation}
		\Gamma_i = -c_i\left(\varphi_i-\varphi_{i,\text{eq}}\right)
	\end{equation}
	Here the $c_i$ are spring constants (a free parameter of the model) and the set $\{\varphi_i\}$ are the angles measured between the $i-1$ and $i+1$ domains. Equally useful is the harmonic energy associated with each torque. These are given by
	\begin{equation}
		U_i = \frac{1}{2}c_i(\varphi_i-\varphi_{i,\text{eq}})^2
	\end{equation} 
	Bead-and-spring models such as this are common in molecular dynamics simulations as a way to model the interactions of high weight molecules.\cite{bead-spring-model}.

@@

	\section{Equations of Motion}
	The dynamics of each domain are determined by equation 3.5 so that in order to simulate the complicated structure we must solve a system of coupled first order differential equations. In doing so, it is convenient to introduce two states to distinguish between dynein's stepping and when both of its binding domains are attached to the microtubule. We shall refer to these configurations hereafter as the one bound state and the both bound state with the idea being that they directly correspond to the pre and post-stroke states defined by the mechanochemical cycle of \cite{cianfrocco2015mechanism}.

@@

		\subsection{The one bound state}
		Figure \ref{fig:one bound} illustrates dynein in the one bound state with new angles $\{\theta_i\}$ chosen for convenience of calculation and computation.\\
		\begin{figure}[!hbt]
			\centering
			\includegraphics[angle=-90,width=0.8\columnwidth]{onebound}
			\caption{The dynein model in the one bound state }
			\label{fig:one bound}  
		\end{figure}

@@

		\subsection{The both bound state}
		For the both bound state, we no longer have a notion of bound and unbound domains. We therefore introduce a new notation, near and far, to identify each dimer. This is not purely for convenience of notation, but rather is important for storing data. We will want the ability to calculate the total step length. Doing this requires knowledge of which motor domain initiates the stepping cycle. \\

		In this state, both binding domains are attached to the microtubule and therefore, a degree of freedom has been removed from our system. This enables us to simplify the dynamic equations in order to take advantage of the new, constrained geometry. To that end, we define a new coordinate system as shown in the following figure.

		\begin{figure}[!hbt]
			\centering
			\includegraphics[width=0.75\columnwidth, angle=-90]{bothbound.pdf}
			\caption{\textbf{Both bound} state with the new angles and lengths shown.}
			\label{fig:bothbound}
		\end{figure}

		Here, $\theta_{nm}$ and $\theta_{fm}$ can be obtained from the one bound state by 
		\begin{equation}
			\theta_{nm} = \theta_1+(\pi-\theta_0) \qquad \theta_{fm}= \theta_3 + (\pi-\theta_4) 
		\end{equation}
		Because the near and far binding domains are fixed, the only domains that we need to time evolve for the both bound state are the two motors (nm, fm) and the tail (T). Therefore, specifying $\theta_{nm}, \theta_{fm}$, and $\ell$ is enough to determine the geometry of the system. The intermediate variables are calculated by wild application of the law of cosines. 
		\begin{align}
			\ell_n &= \sqrt{\ell_s^2+\ell_T^2-2\ell_s\ell_T\cos(2\pi-\theta_{nm})} \\
			\ell_f &=  \sqrt{\ell_s^2+\ell_T^2-2\ell_s\ell_T\cos(2\pi-\theta_{fm})} \\
			\text{}\nonumber\\
			 \theta_n &= \arccos\left(\frac{\ell_n^2+\ell^2-\ell_f^2}{2\ell_n\ell}\right)\\
			\theta_f &= \pi-\arccos\left(\frac{\ell_f^2+\ell^2-\ell_n^2}{2\ell_f\ell}\right)\\
			\text{}\nonumber\\
			\theta_{n'} &= \arccos\left(\frac{\ell_n^2+\ell_s^2-\ell_t^2}{2\ell_n\ell_s}\right) \\
			\theta_{f'} &= \arccos\left(\frac{\ell_f^2+\ell_s^2-\ell_t^2}{2\ell_f\ell_s}\right) \\ 
			\text{}\nonumber \\ 
			\theta_{nb} &\equiv \theta_{n}-\theta_{n'} \\ 
			\theta_{fb} &\equiv \theta_{f}-\theta_{f'} 
		\end{align}
	   	with these angles and lengths defined, we can now easily define the positions 
	   		\begin{align}
	   	{x}'s \quad \begin{cases}
	   	{x}_{nm} = x_n+\ell_s\cos(\theta_{nb}) \\
	   	{x}_{fm} = x_f+\ell_s\cos(\theta_{fb}) \\
	   	{x}_{T} = x_{nm}+\ell_n\cos(\theta_n) 
	   	\end{cases} &
	   	{y}'s \quad \begin{cases}
	   	{y}_{nm} = \ell_s\sin(\theta_{nb}) \\
	   	{y}_{fm} = \ell_s\sin(\theta_{fb}) \\
	   	{y}_{t} = \ell_n\sin(\theta_n) 
	   	\end{cases} 
	   	\end{align}
		Now we have everything we need in order to repeat the process for the one bound state. Differentiating these equations will give us explicit formulas for the time derivative of the coordinates of each domain which we can then compare to the forces acting on each domain via the Brownian motion equation. Fortunately, the effort was well spent as we've reduced (3.14) to a system of six equation with 6 unknowns (one angle and one Lagrange multiplier for each \textit{moving} domain).

@@

	\begin{equation}
	\rho_{ub} = k_{ub}e^{-c(\theta_i-\theta_{i,eq})}
	\end{equation}

### 2019-04-23 papers/waczak-thesis/thesis/thesis.tex
big thesis update 1

	\maketitle

	\chapter*{Abstract}
	test
	\chapter*{Acknowledgments}
	I want to thank ...

>>>>

	\maketitle
	\todo[inline]{TODO: Fix formatting of title page}
	\chapter*{Abstract}
	\vspace{-3 em}word count: \textbf{345}\\

	The dynein molecule is a peculiar motor protein recognized for its odd stepping behavior. Sometimes it steps forwards. Other times it steps backwards. It has even been observed to occasionally shuffle by beginning a step with the same domain many times in row. These unique motions make dynein an interesting object of study, but its status as one of a select few minus-end directed motors makes understanding the mechanics of dynein’s motion vital to our understanding of the cell. When these motors misbehave, fundamental processes like axonal transport and muscle contraction fail. Observational techniques for studying biomolecules have revealed much about the different conformational states achieved by dynein during its stepping cycle and have established a base of relevant statistics such step lengths, velocities, and generated forces among others. Despite these advancements, the ability to image the full molecule for an entire stepping cycle remains beyond the capabilities of the current experimental apparatus.

	To address this gap, we have developed a simplified, two-dimensional model for the dynein molecule. We treat the protein as a system of massive domains held together by rigid rods steered towards equilibrium configurations by spring like torques and time evolved using Brownian dynamics. With our current set of parameters, the model displays directed motion with step length and velocity distributions ( give statistics here ) in agreement with experiment. Our simulations also suggest that there is no correlation between the conformation of dynein before and after a step. Our results indicate that a more computationally efficient simulation can be achieved with a two state system where a post power stroke state is selected from a distribution via a Monte Carlo simulation and binding is then simulated with the full Brownian dynamics. We are optimistic that our simulations will be able to address the power stroke versus winch model debate.

	\todo[inline]{TODO: Add in specific data to results}
	\todo[inline]{TODO: Fix formatting to include "a thesis submitted in completion of..." etc.}

	\chapter*{Acknowledgments}
	I want to thank ...
